---
title: "Affect-AAT model"
author: "Kristina Kobrock"
date: "14 3 2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)
```
```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
# package to extract R data from matlab files
library(R.matlab)

# package for convenience functions (e.g. ggplot2, dplyr, etc.)
library(tidyverse)

# package for Bayesian regression modeling
library(brms)

# package for visualization
library(tidybayes)

# package for plotting (together with ggplot)
library(ggpubr)

# package to visualize 
library(bayesplot)

# package to extract CrIs
library(bayestestR)

# package to extract HDIs
library(HDInterval)

#devtools::install_github("michael-franke/aida-package")
library(aida)

# use the aida-theme for plotting
theme_set(theme_aida())

# global color scheme / non-optimized
# Green RYB: 5EB045, Barn red: 811F0E, Charcoal: 424B54, maximum yellow red: FFBA49, copper crayola: DE8F6E
green <- "#5EB045"
red <- "#811F0E"
gray <- "#424B54"
yellow <- "#FFBA49"
copper <- "#DE8F6E"
black <- "#000000"
project_colors = c(red, green, gray, yellow, copper, black)

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
} 

```

# The role of affect in the Automatic Approach Bias: An empirical investigation.

## The effect of mood induction in the AAB
Load the preprocessed data (see DataAnalysis.Rmd for details).
```{r load-data}
load("affect-aat.Rdata")
```

```{r backtransform-function, echo=FALSE}
backtransform_RT <- function(list, df=df_aat_log1) {
  round(exp(list * attr(df$log_RT_s, 'scaled:scale') + attr(df$log_RT_s, 'scaled:center')),2)
}
```


### AAB model

Fitting a linear model with a gaussian distribution.
```{r data-log-model, echo=FALSE}
# contrast-coding independent variables
df_aat_log1 <- df_aat_log %>% 
  mutate(mood_induction = factor(ifelse(mood_induction == "p", "positive", "negative")),
         condition = factor(condition)) %>% 
  rename(stimulus_valence = stimulus) %>% 
  select(subject_id, trial, condition, mood_induction, log_RT_s, pictures, stimulus_valence, response)
head(df_aat_log1)
```
I chose a weakly informative priors for the model-specific parameter "sigma", and the group-level effects: student_t distribution with 3 degrees of freedom, centered on 0, ranging between -10 and 10 with most values being close to zero. I decided for a slightly more constraining prior on the intercept: a student_t distribution with 5 degrees of freedom, centered on -0.2 ranging from -5 to 5. And an agnostic prior for the population-level effects: normal distribution centered on 0 with wide standard deviation of 10.

```{r priors, echo=FALSE}
priors <- c(set_prior("student_t(3, 0, 2)", class = "Intercept"), # centered on zero, ranges between -10 and 10, most values being closer to zero
            set_prior("student_t(3, 0, 2)", class = "sigma"),
            set_prior("normal(0, 10)", class = "b"), # agnostic prior on fixed effects
            set_prior("student_t(3, 0, 2)", class = "sd") # weakly informative prior on random effects
            ) 
```

```{r constrpriors, echo=FALSE}
constr_priors <- c(set_prior("student_t(5, -0.2, 1)", class = "Intercept"), # constraining prior informed by previous model fit and data: centered on -0.2, ranges between -5 and 5, most values being closer to mu
            set_prior("student_t(3, 0, 2)", class = "sigma"),
            set_prior("normal(0, 10)", class = "b"), # agnostic prior on fixed effects
            set_prior("student_t(3, 0, 2)", class = "sd") # weakly informative prior on random effects
            ) 
```
```{r AAB-model, echo=FALSE}

fit_AABmodel <- brm(log_RT_s ~ condition
                + (condition || subject_id), # I expect differences between individual subjects with regards to condition (correlation not needed), but not with stimulus_valence
                 data = df_aat_log1,
                 family = gaussian(), # model fit is much better with skew_normal() than a normal gaussian() but doesn't converge
                 prior = constr_priors,     
                 iter = 10000,
                 save_pars = save_pars(all=TRUE), # saving pars for later use of Bayes Factors
                 cores = getOption("mc.cores", 2), # can be set to number of cores the hardware possesses (in my case 2) up to the number of chains (4 by default)
                 seed = 1403,
                 file = "MI_model"
                 )
```
```{r summary-mi-model, echo=FALSE}
summary(fit_AABmodel, priors=TRUE)
```

```{r pp-check-mi-model, echo=FALSE}
pp_check(fit_MImodel, ndraws = 50, type = "dens_overlay")
```

The model fit is okay, but it's also very obvious that the normal link function failed to capture the skewed distribution of the data. But due to previous studies that successfully used a linear model with normal link function, the fit is considered good enough.
```{r plot-hdi-posteriors, echo=FALSE}
# Extract the posteriors
posteriors <- fit_MImodel %>%
  spread_draws(b_Intercept, 
               b_conditionincongruent,
               b_mood_inductionpositive,
               `b_conditionincongruent:mood_inductionpositive`) %>%
  # calculate posteriors for each individual level
  mutate(congruent_negative = b_Intercept,
         incongruent_negative = b_Intercept + b_conditionincongruent,
         congruent_positive = b_Intercept + b_mood_inductionpositive,
         incongruent_positive = b_Intercept + b_conditionincongruent + b_mood_inductionpositive 
                                + `b_conditionincongruent:mood_inductionpositive`) %>% 
  select(congruent_negative, incongruent_negative, congruent_positive, incongruent_positive) %>%
  gather(key = "parameter", value = "posterior") %>% 
  group_by(parameter) %>% 
  summarise(mean_posterior = mean(posterior),
            `95lowerCrI` = HDInterval::hdi(posterior, credMass = 0.95)[1],
            `95higherCrI` = HDInterval::hdi(posterior, credMass = 0.95)[2])

posteriors <- posteriors %>% 
  mutate(condition = ifelse(parameter == "congruent_negative" | parameter == "congruent_positive",
                            "congruent", "incongruent"),
         mood_induction = ifelse(parameter == "congruent_negative" | parameter == "incongruent_negative",
                                 "negative", "positive")
  ) %>% 
  select(mood_induction, condition, mean_posterior, `95lowerCrI`, `95higherCrI`) 

# plot
ggplot(data = posteriors, 
       aes(x = condition, y = mean_posterior,
           color = mood_induction, fill = mood_induction)) + 
  geom_errorbar(aes(ymin = `95lowerCrI`, ymax = `95higherCrI`),
                width = 0.2, color = "grey") +
  geom_line(aes(group = 1)) +
  geom_point(size = 5) +
  facet_grid(~mood_induction) +
  guides(color = "none", fill = "none") +
  labs(x = "condition by mood induction",
       y = "posterior log(RT) (scaled)",
       title = "Posterior estimates with credible intervals")
```

Posterior estimates match the data very well. The posterior means are nearly the same as the means from the data distribution (compare to figure above).
```{r extract-posteriors, echo=FALSE}
posteriors_MImodel <- fit_MImodel %>%
  # just taking the population level effects here
  spread_draws(b_Intercept, b_conditionincongruent,
               b_mood_inductionpositive, `b_conditionincongruent:mood_inductionpositive`) %>% 
  mutate(
    congruent_negative = b_Intercept,
    incongruent_negative = b_Intercept + b_conditionincongruent,
    congruent_positive = b_Intercept + b_mood_inductionpositive,
    incongruent_positive = b_Intercept + b_conditionincongruent + 
      b_mood_inductionpositive + `b_conditionincongruent:mood_inductionpositive` 
    ) %>% 
  dplyr::select(congruent_negative, incongruent_negative,
                congruent_positive, incongruent_positive) %>%
  gather(key = "parameter", value = "posterior")

```

```{r differences, echo=FALSE}

# difference between congruent and incongruent condition regardless of mood induction:

# congruent condition:
post_congruent <- posteriors_MImodel %>% 
  filter(parameter == "congruent_negative" | parameter == "congruent_positive") %>% 
  select(posterior) 

# incongruent condition:
post_incongruent <- posteriors_MImodel %>% 
  filter(parameter == "incongruent_negative" | parameter == "incongruent_positive") %>% 
  select(posterior) 

diff_condition <- post_incongruent - post_congruent # positive when RTs in incongruent cond are longer than congruent (as expected)


# negative mood induction:
post_congruent_negMI <- posteriors_MImodel %>% 
  filter(parameter == "congruent_negative") %>% 
  select(posterior) 

post_incongruent_negMI <- posteriors_MImodel %>% 
  filter(parameter == "incongruent_negative") %>% 
  select(posterior)

diff_condition_negMI <- post_incongruent_negMI - post_congruent_negMI # positive when RTs in incongruent condition longer

# positive mood induction:
post_congruent_posMI <- posteriors_MImodel %>% 
  filter(parameter == "congruent_positive") %>% 
  select(posterior)

post_incongruent_posMI <- posteriors_MImodel %>% 
  filter(parameter == "incongruent_positive") %>% 
  select(posterior) 

diff_condition_posMI <- post_incongruent_posMI - post_congruent_posMI # positive when RTs in incongruent cond are longer than congruent (as expected)

```
```{r difference-plot, echo=FALSE}
plot_1 <- diff_condition %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = yellow, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "condition posterior difference"
  ) +
  ylim(0, 13)

plot_2 <- diff_condition_negMI %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = red, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "after negative mood ind."
  ) +
  ylim(0, 13)

plot_3 <- diff_condition_posMI %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = green, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "after positive mood ind."
  ) +
  ylim(0, 13)

ggarrange(plot_1, ggarrange(plot_2, plot_3, ncol=2), nrow=2)
```

```{r difference-in-ms, echo=FALSE}
# transform all posterior estimates to the ms scale:
post_congruent_ms <- backtransform_RT(post_congruent)
post_incongruent_ms <- backtransform_RT(post_incongruent)
diff_condition_ms <- post_incongruent_ms - post_congruent_ms

post_incongruent_negMI_ms <- backtransform_RT(post_incongruent_negMI)
post_congruent_negMI_ms <- backtransform_RT(post_congruent_negMI)
diff_condition_negMI_ms <- post_incongruent_negMI_ms - post_congruent_negMI_ms

post_incongruent_posMI_ms <- backtransform_RT(post_incongruent_posMI)
post_congruent_posMI_ms <- backtransform_RT(post_congruent_posMI)
diff_condition_posMI_ms <- post_incongruent_posMI_ms - post_congruent_posMI_ms

```
```{r plot-differences-ms, echo=FALSE}
plot_1 <- diff_condition_ms %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = yellow, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "condition posterior difference [ms]"
  ) +
  ylim(0, 0.05)

plot_2 <- diff_condition_negMI_ms %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = red, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "after negative mood ind."
  ) +
  ylim(0, 0.05)

plot_3 <- diff_condition_posMI_ms %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = green, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "after positive mood ind."
  ) +
  ylim(0, 0.05)

ggarrange(plot_1, ggarrange(plot_2, plot_3, ncol=2), nrow=2)
```
```{r describe-differences-ms, echo=FALSE}

df_diffcondition <- data.frame(
  posteriors = "condition difference",
  median = median(diff_condition_ms$posterior),
  CrI_lower = HDInterval::hdi(diff_condition_ms$posterior)[1],
  CrI_upper = HDInterval::hdi(diff_condition_ms$posterior)[2]
)
df_diffnegMI <- data.frame(
  posteriors = "difference after negative mood induction",
  median = median(diff_condition_negMI_ms$posterior),
  CrI_lower = HDInterval::hdi(diff_condition_negMI_ms$posterior)[1],
  CrI_upper = HDInterval::hdi(diff_condition_negMI_ms$posterior)[2]
)
df_diffposMI <- data.frame(
  posteriors = "difference after positive mood induction",
  median = median(diff_condition_posMI_ms$posterior),
  CrI_lower = HDInterval::hdi(diff_condition_posMI_ms$posterior)[1],
  CrI_upper = HDInterval::hdi(diff_condition_posMI_ms$posterior)[2]
)

differences_in_ms <- tibble(rbind(
  df_diffcondition,
  df_diffnegMI,
  df_diffposMI
))

differences_in_ms
```

```{r bayestests, echo=FALSE}
MImodel_rope_range = rope_range(fit_MImodel) # calculates the rope range using the tenth (1/10 = 0.1) of the standard deviation (SD) of the response variable (panas_score), which can be considered as a “negligible” effect size (Cohen, 1988)

print("condition difference regardless of mood induction:")
posterior_diffcondition <- describe_posterior(diff_condition, centrality = "median", ci = 0.95, ci_method = "hdi", rope_range = MImodel_rope_range, rope_ci = 0.95) %>% print()

print("condition difference after negative mood induction:")
posterior_diffc_negMI <- describe_posterior(diff_condition_negMI, rope_range = MImodel_rope_range) %>% print()

print("condition difference after positive mood induction:")
posterior_diffc_posMI <- describe_posterior(diff_condition_posMI, rope_range = MImodel_rope_range) %>% print()

print("Is there a difference between the AAB (i.e. difference in RTs between conditions) after positive compared to after negative MI?")
print("positive AAB in negative AAB:")
posterior_diff_AAB <- describe_posterior(diff_condition_posMI, centrality = "none", rope_range = c(posterior_diffc_negMI$CI_low, posterior_diffc_negMI$CI_high), test = c("rope"))
print(posterior_diff_AAB)

print("negative AAB in positive AAB:")
posterior_diff_AAB2 <- describe_posterior(diff_condition_negMI, centrality = "none", rope_range = c(posterior_diffc_posMI$CI_low, posterior_diffc_posMI$CI_high), test = c("rope"))
print(posterior_diff_AAB2)

```

The hypothesis tests according to the hypotheses above support the results:
```{r hypothesis-tests, echo=TRUE}

# Hypothesis 1: After negative mood induction, reaction times are shorter in the incongruent than in the congruent condition.
hypothesis(fit_MImodel, "conditionincongruent < 0")
# --> reject hypothesis, test other way:
hypothesis(fit_MImodel, "conditionincongruent > 0")

# Hypothesis 2: After positive mood induction, reaction times are longer in the incongruent than in the congruent condition.
hypothesis(fit_MImodel, "conditionincongruent + `conditionincongruent:mood_inductionpositive` > 0")
```
```{r hypothesistests-MIdifferences, echo=TRUE}
# are reaction times different from each other in the different mood induction conditions?

# Is congruent negative different from congruent positive?
hypothesis(fit_MImodel, "mood_inductionpositive < 0")

# Is incongruent negative different from incongruent positive?
hypothesis(fit_MImodel, "mood_inductionpositive + `conditionincongruent:mood_inductionpositive` < 0")

# ---> not really. there is some evidence (80%) for a difference between mood inductions for incongruent conditions, but it is not substantial
```

