---
title: "DataViz_AAAT"
author: "Kristina Kobrock"
date: "10 1 2022"
output: html_document
---
```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(R.matlab)
library(tidyverse)
library(brms)
```

# The role of affect in the Automatic Approach Bias: An empirical investigation.

## Data Analysis

### Read and preprocess the data
```{r read_data}
files_1stsession = file.path("./data/1stsession",  list.files(path="./data/1stsession", pattern=".mat$"))
files_2ndsession = file.path("./data/2ndsession", list.files(path="./data/2ndsession", pattern=".mat$"))
session_1 <- lapply(files_1stsession, readMat)
session_2 <- lapply(files_2ndsession, readMat)
```
```{r preprocess_function_aat}
preprocess_aat <- function(data) {
  
  subjectindex <- data$subject.index # unique subject ID
  subjectname <- data$subject.name
  subjectage <- data$subject.age
  subjectgender <- data$subject.gender
  subjecthand <- data$subject.hand
  date <- data$subject.date
  session_nr <- data$subject.session
  mood_induction <- data$mood.induction

  block_order <- data$category # block order: a=congruent-incongruent, b=incongruent-congruent
  if (block_order == 'a') {
    block <- c(rep(c("congruent", "incongruent", "congruent", "incongruent"), times=c(44,44,44,44)))  
  }
  else {
    block <- c(rep(c("incongruent", "congruent", "incongruent", "congruent"), times=c(44,44,44,44)))
  }
  
  # sequences: picture order (number 1-44 are positive, 45-88 are negative   pictures)
  sequence_1 <- c()
  sequence_2 <- c()
  for (i in 1:176){
    if ((i %% 2) != 0){
      sequence_1 <- c(sequence_1, data$sequences[[i]])
    }
    else {
      sequence_2 <- c(sequence_2, data$sequences[[i]])
    }
  }
  pushtimes <- data$pushtimes # only push trials
  pulltimes <- data$pulltimes # only pull trials
  alltimes <- data$alltimes # all reaction times


  df <- data.frame(subjectindex = rep(subjectindex, 176),
                   subjectname = rep(subjectname, 176),
                   subjectgender = rep(subjectgender, 176),
                   subjectage = rep(subjectage, 176),
                   subjecthand = rep(subjecthand, 176),
                   session_nr = rep(session_nr, 176),
                   mood_induction = rep(mood_induction, 176),
                   block_order = rep(block_order, 176),
                   trial = c(1:176),
                   block = block,
                   pictures = c(sequence_1, sequence_2),
                   response = pushtimes[1:176],
                   pushtimes = pushtimes[1:176],
                   pulltimes = pulltimes[1:176],
                   reaction_times = alltimes[1:176]) %>% 
    mutate(stimulus = ifelse(pictures <= 44, "positive", "negative"),
           response = ifelse(response == 0, "pull", "push"),
           correct = ifelse(((block == "congruent" & stimulus == "positive" & response == "pull") |
                               (block == "congruent" & stimulus == "negative" & response == "push") |
                               (block == "incongruent" & stimulus == "positive" & response == "push") |
                               (block == "incongruent" & stimulus == "negative" & response == "pull")), 1, 0))
}
```
```{r peprocess_function_affect}

preprocess_affect <- function(data) {
  
  subjectindex <- data$subject.index # unique subject ID
  subjectname <- data$subject.name
  subjectage <- data$subject.age
  subjectgender <- data$subject.gender
  subjecthand <- data$subject.hand
  date <- data$subject.date
  session_nr <- data$subject.session
  mood_induction <- data$mood.induction
  panas1_item1 <- data$a1item.1
  panas1_item2 <- data$a1item.2
  panas1_item3 <- data$a1item.3
  panas1_item4 <- data$a1item.4
  panas1_item5 <- data$a1item.5
  panas1_item6 <- data$a1item.6
  panas1_item7 <- data$a1item.7
  panas1_item8 <- data$a1item.8
  panas1_item9 <- data$a1item.9
  panas1_item10 <- data$a1item.10
  panas1_item11 <- data$a1item.11
  panas1_item12 <- data$a1item.12
  panas1_item13 <- data$a1item.13
  panas1_item14 <- data$a1item.14
  panas1_item15 <- data$a1item.15
  panas1_item16 <- data$a1item.16
  panas1_item17 <- data$a1item.17
  panas1_item18 <- data$a1item.18
  panas1_item19 <- data$a1item.19
  panas1_item20 <- data$a1item.20
  panas2_item1 <- data$a2item.1
  panas2_item2 <- data$a2item.2
  panas2_item3 <- data$a2item.3
  panas2_item4 <- data$a2item.4
  panas2_item5 <- data$a2item.5
  panas2_item6 <- data$a2item.6
  panas2_item7 <- data$a2item.7
  panas2_item8 <- data$a2item.8
  panas2_item9 <- data$a2item.9
  panas2_item10 <- data$a2item.10
  panas2_item11 <- data$a2item.11
  panas2_item12 <- data$a2item.12
  panas2_item13 <- data$a2item.13
  panas2_item14 <- data$a2item.14
  panas2_item15 <- data$a2item.15
  panas2_item16 <- data$a2item.16
  panas2_item17 <- data$a2item.17
  panas2_item18 <- data$a2item.18
  panas2_item19 <- data$a2item.19
  panas2_item20 <- data$a2item.20
  
  panas_1 <- c(panas1_item1, panas1_item2, panas1_item3, panas1_item4, panas1_item5, panas1_item6, panas1_item7, panas1_item8, panas1_item9, panas1_item10, panas1_item11, panas1_item12, panas1_item13, panas1_item14, panas1_item15, panas1_item16, panas1_item17, panas1_item18, panas1_item19, panas1_item20)
  
  panas_2 <- c(panas2_item1, panas2_item2, panas2_item3, panas2_item4, panas2_item5, panas2_item6, panas2_item7, panas2_item8, panas2_item9, panas2_item10, panas2_item11, panas2_item12, panas2_item13, panas2_item14, panas2_item15, panas2_item16, panas2_item17, panas2_item18, panas2_item19, panas2_item20)
  
panas_item = c("interested", "distressed", "excited", "upset", "strong", "guilty", "scared", "hostile", "enthusiastic", "proud", "irritable", "alert", "ashamed", "inspired", "nervous", "determined", "attentive", "jittery", "active", "afraid")

item_affect = c("positive", "negative", "positive", "negative", "positive", "negative", "negative", "negative", "positive", "positive", "negative", "positive", "negative", "positive", "negative", "positive", "positive", "negative", "positive", "negative")

df <- data.frame(subjectindex = rep(subjectindex, 20),
                 subjectname = rep(subjectname, 20),
                 subjectgender = rep(subjectgender, 20),
                 subjectage = rep(subjectage, 20),
                 subjecthand = rep(subjecthand, 20),
                 session_nr = rep(session_nr, 20),
                 mood_induction = rep(mood_induction, 20),
                 panas_item = panas_item,
                 item_nr = c(1:20),
                 item_affect = item_affect,
                 panas_1 = as.numeric(panas_1),
                 panas_2 = as.numeric(panas_2)) %>% 
  mutate(mood_induction = ifelse(mood_induction == "p", "positive", "negative")) %>% 
  rename('1' = panas_1, '2' = panas_2) %>% 
  pivot_longer(cols = c('1', '2'), names_to = "panas_nr", values_to = "item_score")
}
```
```{r preprocessing}
df_aat <- data.frame()
df_affect <- data.frame()
for (data_file in session_1) {
  df_single <- data.frame(data_file$subjectdata)[[1]]
  df_aat <- rbind(df_aat, preprocess_aat(df_single))
  df_affect <- rbind(df_affect, preprocess_affect(df_single))
}
for (data_file in session_2) {
  df_single <- data.frame(data_file$subjectdata)[[1]]
  df_aat <- rbind(df_aat, preprocess_aat(df_single))
  df_affect <- rbind(df_affect, preprocess_affect(df_single))
}
head(df_aat)
head(df_affect)
```
```{r}
# How many data points do I have by now?
# rough proxy:
# print(unique(df_aat$subjectindex)) # 76

# check whether for each subjectindex is data available from two sessions and two mood inductions (and the same block order)
df_aat_check <- df_aat %>% 
  group_by(subjectindex, session_nr, mood_induction, block_order) %>% 
  summarize() %>% 
  print

# data from 1st session subject 8 missing (nb-pb) --> exc subject 8
# exclude subject 10 because she didn't fulfill the criteria (pb-nb) --> exc subject 10
# second session from subject 16 missing (na-pa) --> exc subject 16
# second session from subject 24 missing (nb-pb) 
# subject 33: mistakenly assigned na-pb instead of na-pa --> exc subject 33
# second session from subject 50 missing (pb-nb) 
# ...

# only keep data when data from both sessions is available and no measurement errors occured
df_aat_excs <- df_aat %>% 
  filter(subjectindex != 8 
         & subjectindex != 10
         & subjectindex != 16
         & subjectindex != 33)

df_affect_excs <- df_affect %>% 
  filter(subjectindex != 8 
         & subjectindex != 10
         & subjectindex != 16
         & subjectindex != 33)

```

### Mood induction
Is the mood induction successful? 
--> positive mood induction: panas_scale more positive in panas_2 than in panas_1
--> negative mood induction: panas_scale more negative in panas_2 than in panas_1
1. calculate overall panas_scales for each participant
2. compare and check for statistical significance

```{r calc_panas_scores}
df_affect_scores <- df_affect_exc %>% 
  group_by(subjectindex, session_nr, mood_induction, panas_nr, item_affect) %>% 
  summarise(panas_score = sum(item_score)) %>% 
  ungroup()
head(df_affect_scores)
```
#### Visual inspection:
```{r compare_scores}
df_affect_agg <- df_affect_scores %>% 
  group_by(mood_induction, panas_nr, item_affect) %>% 
  summarise(mean = mean(panas_score)) %>% 
  ungroup()
df_affect_agg
```
```{r plot_mood}
df_affect_scores %>% 
  ggplot(aes(x = panas_nr, y = panas_score, color = mood_induction)) + 
  facet_wrap(item_affect ~ mood_induction) + 
  geom_point(alpha = 0.1, size = 1,
             # adding a little bit of jitter to make the points visible
             position = position_jitter(width = 0.1)) +
  # adding means and a connecting line in order to make the differences between locations and conditions 
  # visible
  geom_point(data = df_affect_agg,
             aes(x = panas_nr, y = mean, fill = mood_induction), size = 4) +
  geom_line(data = df_affect_agg,
            aes(x = panas_nr, y = mean, group = mood_induction))
```

### AAT data
```{r plot_raw_data, echo=FALSE}
df_aat_excs %>% 
  ggplot(aes(x = reaction_times)) +
  geom_histogram(binwidth = 0.1, fill="lightblue", color="blue") + 
  labs(
    title = "Raw data",
    x = "reaction times (s)"
  )
```
```{r exclude}
# exclude practice trials (first 4 of each block)
df_aat_exct <- df_aat_excs %>% 
  filter(trial > 4
         & (trial < 45 | trial > 48)
         & (trial < 89 | trial > 92)
         & (trial < 133 | trial > 136))
df_aat_raw <- df_aat_exct
nr_raw <- nrow(df_aat_raw) # %>% print # 23360
df_aat_exc <- df_aat_raw %>% 
  filter(correct == "1" & reaction_times > 0.15) #%>%  # only correct responses and reaction times above 150ms will be analysed
nr_exc <- nrow(df_aat_exc) # %>% print # 22449
data_exc <- round((1 - nr_exc / nr_raw)*100, 2) # %>% print # 3.9 %
```
```{r log-transform}
# standardize and log-transform reaction times
df_aat_log <- df_aat_exc %>% 
  mutate(log_RT = log(reaction_times))
df_aat_log$log_RT_s <- scale(df_aat_log$log_RT, scale = TRUE)
```
```{r plot2_log_data}
df_aat_log %>% 
  ggplot(aes(x = log_RT_s)) +
  geom_histogram(binwidth = 0.1, fill="lightblue", color="blue") + 
  labs(
    title = "Log-transformed data",
    x = "log(RT)"
  )
```
```{r}
# creating an aggregated data frame grouped by mood_induction and block with the respective means
df_aat_agg1 <- df_aat_log %>% 
  group_by(mood_induction, block) %>% 
  summarise(mean = mean(log_RT_s))

# plot
df_aat_log %>% 
  # plotting both mood_induction and block using facet_wrap
  ggplot(aes(x = block, y = log_RT_s, color = mood_induction)) +
  facet_wrap(~ mood_induction) +
  geom_point(alpha = 0.1, size = 1,
             # adding a little bit of jitter to make the points visible
             position = position_jitter(width = 0.1)) +
  # adding means and a connecting line in order to make the differences between locations and conditions 
  # visible
  geom_point(data = df_aat_agg1,
             aes(x = block, y = mean, fill = mood_induction), size = 4) +
  geom_line(data = df_aat_agg1,
            aes(x = block, y = mean, group = mood_induction)) +
  # custom colours
  scale_colour_manual(values = c("green4", "red2")) +
  # axes labels and titles
  labs(
    #title = "Effects of mood induction and block (congruent/incongruent)",
    #subtitle = "Hypotheses 1 & 2",
    y = "log(RT)"
  )
```

```{r}
log_model <- brm(log_RT_s ~ block * mood_induction, 
                data = df_aat_log)
```

```{r}
log_model
```
```{r}
pp_check(log_model)
```

```{r}
log_model_aat <- brm(log_RT_s ~ block,
                 data = df_aat_log)
```
```{r}
pp_check(log_model_aat)
```

```{r}
log_model_aat
```

```{r}
#model_max_re <- brm(reaction_times ~ block * mood_induction +
#                      (1 | trial) +
#                      (1 | subjectindex) + 
#                      (1 | stimulus) + 
#                      (1 | response),
#                    data = df_aat_prep)
```

```{r}
#model_max_re
```

```{r}
# model without transformed reaction times

nolog_model <- brm(reaction_times ~ block * mood_induction, 
                   family = Gamma(log),
                   data = df_aat_exc)
```
```{r}
nolog_model
```

```{r}
pp_check(nolog_model)
pp_check(log_model)
```

