---
title: "Mood induction item analysis"
author: "Kristina Kobrock"
date: "14 3 2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)
```
```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
# package to extract R data from matlab files
library(R.matlab)

# package for convenience functions (e.g. ggplot2, dplyr, etc.)
library(tidyverse)

# package for Bayesian regression modeling
library(brms)

# package for visualization
library(tidybayes)

# package for plotting (together with ggplot)
library(ggpubr)

# package to visualize 
library(bayesplot)

# package to extract CrIs
library(bayestestR)

# package to extract HDIs
library(HDInterval)

#devtools::install_github("michael-franke/aida-package")
library(aida)

# use the aida-theme for plotting
theme_set(theme_aida())

# global color scheme / non-optimized
# Green RYB: 5EB045, Barn red: 811F0E, Charcoal: 424B54, maximum yellow red: FFBA49, copper crayola: DE8F6E
green <- "#5EB045"
red <- "#811F0E"
gray <- "#424B54"
yellow <- "#FFBA49"
copper <- "#DE8F6E"
black <- "#000000"
project_colors = c(red, green, gray, yellow, copper, black)

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
} 

```

# The role of affect in the Automatic Approach Bias: An empirical investigation.

## Mood induction
Load the preprocessed data (see DataAnalysis.Rmd for details).
```{r load-data}
load("panas.Rdata")
```

### Investigating the role of specific PANAS questionnaire items on the outcome
Split data into two different datasets: one for PA and one for NA
```{r split-data}
NA_data <- df_affect_mutate %>% 
  filter(item_affect == "NA") 

PA_data <- df_affect_mutate %>% 
  filter(item_affect == "PA")
```
Run two different models to predict PA and NA scales, respectively:
```{r na-model, echo=FALSE}
NA_model <- brm(panas_score ~ panas_item,
                  iter = 2000,
                  data = NA_data,
                  seed = 603,
                  file = "NA_model_onlyitem")
```
```{r pa-model, echo=FALSE}
PA_model <- brm(panas_score ~ panas_item,
                  iter = 2000,
                  data = PA_data,
                  seed = 603,
                  file = "PA_model_onlyitem")
```
```{r sum-na-model, echo=FALSE}
summary(NA_model)
```
```{r sum-pa-model, echo=FALSE}
summary(PA_model)
```

