---
title: "Mood induction item analysis"
author: "Kristina Kobrock"
date: "14 3 2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)
```
```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
# package to extract R data from matlab files
library(R.matlab)

# package for convenience functions (e.g. ggplot2, dplyr, etc.)
library(tidyverse)

# package for Bayesian regression modeling
library(brms)

# package for visualization
library(tidybayes)

# package for plotting (together with ggplot)
library(ggpubr)

# package to visualize 
library(bayesplot)

# package to extract CrIs
library(bayestestR)

# package to extract HDIs
library(HDInterval)

#devtools::install_github("michael-franke/aida-package")
library(aida)

# use the aida-theme for plotting
theme_set(theme_aida())

# global color scheme / non-optimized
# Green RYB: 5EB045, Barn red: 811F0E, Charcoal: 424B54, maximum yellow red: FFBA49, copper crayola: DE8F6E
green <- "#5EB045"
red <- "#811F0E"
gray <- "#424B54"
yellow <- "#FFBA49"
copper <- "#DE8F6E"
black <- "#000000"
project_colors = c(red, green, gray, yellow, copper, black)

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
} 

```

# The role of affect in the Automatic Approach Bias: An empirical investigation.

## Mood induction
Load the preprocessed data (see DataAnalysis.Rmd for details).
```{r load-data}
load("panas_item.Rdata")
```

### Investigating the role of specific PANAS questionnaire items on the outcome
```{r prepare-data}
df_affect_scores <- df_affect_excs %>% 
  group_by(subject_id, session_nr, mood_induction, subjectgender, panas_nr, item_affect) %>% 
  mutate(panas_score = sum(item_score)) %>% 
  ungroup() %>% 
  select(subject_id, session_nr, mood_induction, panas_nr, item_affect, panas_item, item_nr, item_score, panas_score)

df_affect_mutate <- df_affect_scores %>% 
  mutate(panas_score = ifelse(item_affect == "NA", -panas_score, panas_score),
         panas_item = factor(panas_item),
         mood_induction = factor(mood_induction),
         panas_nr = factor(panas_nr))
  
```
Split data into two different datasets: one for PA and one for NA
```{r split-data}
NA_data <- df_affect_mutate %>% 
  filter(item_affect == "NA"
         & mood_induction == "positive")

PA_data <- df_affect_mutate %>% 
  filter(item_affect == "PA"
         & mood_induction == "negative")

```

Run two different models to predict PA and NA scales, respectively:
```{r na-model, echo=FALSE}
NA_model <- brm(panas_score ~ panas_item * panas_nr + item_score,
                  iter = 2000,
                  data = NA_data,
                  seed = 2203,
                  file = "NA_model")
```
```{r sum-na-model, echo=FALSE}
summary(NA_model)
```
```{r extract-posteriors-na, echo=FALSE}
posteriors_NA <- NA_model %>% 
  spread_draws(b_Intercept, b_panas_itemashamed, b_panas_itemdistressed, b_panas_itemguilty, 
               b_panas_itemhostile, b_panas_itemirritable, b_panas_itemjittery, b_panas_itemnervous, 
               b_panas_itemscared, b_panas_itemupset, b_panas_nr2, b_item_score,
               `b_panas_itemashamed:panas_nr2`, `b_panas_itemdistressed:panas_nr2`,
               `b_panas_itemguilty:panas_nr2`, `b_panas_itemhostile:panas_nr2`,
               `b_panas_itemirritable:panas_nr2`, `b_panas_itemjittery:panas_nr2`,
               `b_panas_itemnervous:panas_nr2`, `b_panas_itemscared:panas_nr2`,
               `b_panas_itemupset:panas_nr2`) %>% 
  mutate(afraid_before = b_Intercept, afraid_after = b_Intercept + b_panas_nr2,
         ashamed_before = b_Intercept + b_panas_itemashamed, ashamed_after = b_Intercept + b_panas_itemashamed + b_panas_nr2 + `b_panas_itemashamed:panas_nr2`,
         distressed_before = b_Intercept + b_panas_itemdistressed, distressed_after = b_Intercept + b_panas_itemdistressed + b_panas_nr2 + `b_panas_itemdistressed:panas_nr2`,
         guilty_before = b_Intercept + b_panas_itemguilty, guilty_after = b_Intercept + b_panas_itemguilty + b_panas_nr2 + `b_panas_itemguilty:panas_nr2`,
         hostile_before = b_Intercept + b_panas_itemhostile, hostile_after = b_Intercept + b_panas_itemhostile + b_panas_nr2 + `b_panas_itemhostile:panas_nr2`,
         irritable_before = b_Intercept + b_panas_itemirritable, irritable_after = b_Intercept + b_panas_itemirritable + b_panas_nr2 + `b_panas_itemirritable:panas_nr2`,
         jittery_before = b_Intercept + b_panas_itemjittery, jittery_after = b_Intercept + b_panas_itemjittery + b_panas_nr2 + `b_panas_itemjittery:panas_nr2`,
         nervous_before = b_Intercept + b_panas_itemnervous, nervous_after = b_Intercept + b_panas_itemnervous + b_panas_nr2 + `b_panas_itemnervous:panas_nr2`,
         scared_before = b_Intercept + b_panas_itemscared, scared_after = b_Intercept + b_panas_itemscared + b_panas_nr2 + `b_panas_itemscared:panas_nr2`,
         upset_before = b_Intercept + b_panas_itemupset, upset_after = b_Intercept + b_panas_itemupset + b_panas_nr2 + `b_panas_itemupset:panas_nr2`) %>% 
  select(afraid_before, afraid_after, ashamed_before, ashamed_after,
         distressed_before, distressed_after, guilty_before, guilty_after,
         hostile_before, hostile_after, irritable_before, irritable_after,
         jittery_before, jittery_after, nervous_before, nervous_after,
         scared_before, scared_after, upset_before, upset_after) %>% 
  gather(key = "parameter", value = "posterior")
```
```{r differences-na, echo=FALSE}
# difference between before and after mood induction for item afraid:

# before:
pos_afraid_before <- posteriors_NA %>% 
  filter(parameter == "afraid_before") %>% 
  select(posterior) 

# after:
pos_afraid_after <- posteriors_NA %>% 
  filter(parameter == "afraid_after") %>% 
  select(posterior) 

diff_afraid <- pos_afraid_after - pos_afraid_before # positive 

# ashamed
# before:
pos_ashamed_before <- posteriors_NA %>% 
  filter(parameter == "ashamed_before") %>% 
  select(posterior) 

# after:
pos_ashamed_after <- posteriors_NA %>% 
  filter(parameter == "ashamed_after") %>% 
  select(posterior) 

diff_ashamed <- pos_ashamed_after - pos_ashamed_before # positive 


# distressed
# before:
pos_distressed_before <- posteriors_NA %>% 
  filter(parameter == "distressed_before") %>% 
  select(posterior) 

# after:
pos_distressed_after <- posteriors_NA %>% 
  filter(parameter == "distressed_after") %>% 
  select(posterior) 

diff_distressed <- pos_distressed_after - pos_distressed_before # positive 


# guilty
# before:
pos_guilty_before <- posteriors_NA %>% 
  filter(parameter == "guilty_before") %>% 
  select(posterior) 

# after:
pos_guilty_after <- posteriors_NA %>% 
  filter(parameter == "guilty_after") %>% 
  select(posterior) 

diff_guilty <- pos_guilty_after - pos_guilty_before # positive 


# hostile
# before:
pos_hostile_before <- posteriors_NA %>% 
  filter(parameter == "hostile_before") %>% 
  select(posterior) 

# after:
pos_hostile_after <- posteriors_NA %>% 
  filter(parameter == "hostile_after") %>% 
  select(posterior) 

diff_hostile <- pos_hostile_after - pos_hostile_before # positive 


# irritable
# before:
pos_irritable_before <- posteriors_NA %>% 
  filter(parameter == "irritable_before") %>% 
  select(posterior) 

# after:
pos_irritable_after <- posteriors_NA %>% 
  filter(parameter == "irritable_after") %>% 
  select(posterior) 

diff_irritable <- pos_irritable_after - pos_irritable_before # positive 


# jittery
# before:
pos_jittery_before <- posteriors_NA %>% 
  filter(parameter == "jittery_before") %>% 
  select(posterior) 

# after:
pos_jittery_after <- posteriors_NA %>% 
  filter(parameter == "jittery_after") %>% 
  select(posterior) 

diff_jittery <- pos_jittery_after - pos_jittery_before # positive 



# nervous
# before:
pos_nervous_before <- posteriors_NA %>% 
  filter(parameter == "nervous_before") %>% 
  select(posterior) 

# after:
pos_nervous_after <- posteriors_NA %>% 
  filter(parameter == "nervous_after") %>% 
  select(posterior) 

diff_nervous <- pos_nervous_after - pos_nervous_before # positive 


# scared
# before:
pos_scared_before <- posteriors_NA %>% 
  filter(parameter == "scared_before") %>% 
  select(posterior) 

# after:
pos_scared_after <- posteriors_NA %>% 
  filter(parameter == "scared_after") %>% 
  select(posterior) 

diff_scared <- pos_scared_after - pos_scared_before # positive 


# upset
# before:
pos_upset_before <- posteriors_NA %>% 
  filter(parameter == "upset_before") %>% 
  select(posterior) 

# after:
pos_upset_after <- posteriors_NA %>% 
  filter(parameter == "upset_after") %>% 
  select(posterior) 

diff_upset <- pos_upset_after - pos_upset_before # positive 

```
```{r na-posterior-descriptions, echo=FALSE}
rope_range <- rope_range(NA_model)

print("afraid")
describe_posterior(diff_afraid, rope_range = rope_range)
print("ashamed")
describe_posterior(diff_ashamed, rope_range = rope_range)
print("distressed")
describe_posterior(diff_distressed, rope_range = rope_range)
print("guilty")
describe_posterior(diff_guilty, rope_range = rope_range)
print("hostile")
describe_posterior(diff_hostile, rope_range = rope_range)
print("irritable")
describe_posterior(diff_irritable, rope_range = rope_range)
print("jittery")
describe_posterior(diff_jittery, rope_range = rope_range)
print("nervous")
describe_posterior(diff_nervous, rope_range = rope_range)
print("scared")
describe_posterior(diff_scared, rope_range = rope_range)
print("upset")
describe_posterior(diff_upset, rope_range = rope_range)
```


### PA
```{r pa-model, echo=FALSE}
PA_model <- brm(panas_score ~ panas_item * panas_nr + item_score,
                  iter = 2000,
                  data = PA_data,
                  seed = 2203,
                  file = "PA_model")
```
```{r summary-pa, echo=FALSE}
summary(PA_model)
```
```{r extract-posteriors-pa, echo=FALSE}
posteriors_PA <- PA_model %>% 
  spread_draws(b_Intercept, b_panas_itemalert, b_panas_itemattentive, b_panas_itemdetermined, 
               b_panas_itementhusiastic, b_panas_itemexcited, b_panas_iteminspired, b_panas_iteminterested, 
               b_panas_itemproud, b_panas_itemstrong, b_panas_nr2, b_item_score,
               `b_panas_itemalert:panas_nr2`, `b_panas_itemattentive:panas_nr2`,
               `b_panas_itemdetermined:panas_nr2`, `b_panas_itementhusiastic:panas_nr2`,
               `b_panas_itemexcited:panas_nr2`, `b_panas_iteminspired:panas_nr2`,
               `b_panas_iteminterested:panas_nr2`, `b_panas_itemproud:panas_nr2`,
               `b_panas_itemstrong:panas_nr2`) %>% 
  mutate(active_before = b_Intercept, active_after = b_Intercept + b_panas_nr2,
         alert_before = b_Intercept + b_panas_itemalert, alert_after = b_Intercept + b_panas_itemalert + b_panas_nr2 + `b_panas_itemalert:panas_nr2`,
         attentive_before = b_Intercept + b_panas_itemattentive, attentive_after = b_Intercept + b_panas_itemattentive + b_panas_nr2 + `b_panas_itemattentive:panas_nr2`,
         determined_before = b_Intercept + b_panas_itemdetermined, determined_after = b_Intercept + b_panas_itemdetermined + b_panas_nr2 + `b_panas_itemdetermined:panas_nr2`,
         enthusiastic_before = b_Intercept + b_panas_itementhusiastic, enthusiastic_after = b_Intercept + b_panas_itementhusiastic + b_panas_nr2 + `b_panas_itementhusiastic:panas_nr2`,
         excited_before = b_Intercept + b_panas_itemexcited, excited_after = b_Intercept + b_panas_itemexcited + b_panas_nr2 + `b_panas_itemexcited:panas_nr2`,
         inspired_before = b_Intercept + b_panas_iteminspired, inspired_after = b_Intercept + b_panas_iteminspired + b_panas_nr2 + `b_panas_iteminspired:panas_nr2`,
         interested_before = b_Intercept + b_panas_iteminterested, interested_after = b_Intercept + b_panas_iteminterested + b_panas_nr2 + `b_panas_iteminterested:panas_nr2`,
         proud_before = b_Intercept + b_panas_itemproud, proud_after = b_Intercept + b_panas_itemproud + b_panas_nr2 + `b_panas_itemproud:panas_nr2`,
         strong_before = b_Intercept + b_panas_itemstrong, strong_after = b_Intercept + b_panas_itemstrong + b_panas_nr2 + `b_panas_itemstrong:panas_nr2`) %>% 
  select(active_before, active_after, alert_before, alert_after,
         attentive_before, attentive_after, determined_before, determined_after,
         enthusiastic_before, enthusiastic_after, excited_before, excited_after,
         inspired_before, inspired_after, interested_before, interested_after,
         proud_before, proud_after, strong_before, strong_after) %>% 
  gather(key = "parameter", value = "posterior")
```
```{r differences-pa, echo=FALSE}
# difference between before and after mood induction for item active:

# before:
neg_active_before <- posteriors_PA %>% 
  filter(parameter == "active_before") %>% 
  select(posterior) 

# after:
neg_active_after <- posteriors_PA %>% 
  filter(parameter == "active_after") %>% 
  select(posterior) 

diff_active <- neg_active_after - neg_active_before # positive 

# alert
# before:
neg_alert_before <- posteriors_PA %>% 
  filter(parameter == "alert_before") %>% 
  select(posterior) 

# after:
neg_alert_after <- posteriors_PA %>% 
  filter(parameter == "alert_after") %>% 
  select(posterior) 

diff_alert <- neg_alert_after - neg_alert_before # positive 


# attentive
# before:
neg_attentive_before <- posteriors_PA %>% 
  filter(parameter == "attentive_before") %>% 
  select(posterior) 

# after:
neg_attentive_after <- posteriors_PA %>% 
  filter(parameter == "attentive_after") %>% 
  select(posterior) 

diff_attentive <- neg_attentive_after - neg_attentive_before # positive 


# determined
# before:
neg_determined_before <- posteriors_PA %>% 
  filter(parameter == "determined_before") %>% 
  select(posterior) 

# after:
neg_determined_after <- posteriors_PA %>% 
  filter(parameter == "determined_after") %>% 
  select(posterior) 

diff_determined <- neg_determined_after - neg_determined_before # positive 


# enthusiastic
# before:
neg_enthusiastic_before <- posteriors_PA %>% 
  filter(parameter == "enthusiastic_before") %>% 
  select(posterior) 

# after:
neg_enthusiastic_after <- posteriors_PA %>% 
  filter(parameter == "enthusiastic_after") %>% 
  select(posterior) 

diff_enthusiastic <- neg_enthusiastic_after - neg_enthusiastic_before # positive 


# excited
# before:
neg_excited_before <- posteriors_PA %>% 
  filter(parameter == "excited_before") %>% 
  select(posterior) 

# after:
neg_excited_after <- posteriors_PA %>% 
  filter(parameter == "excited_after") %>% 
  select(posterior) 

diff_excited <- neg_excited_after - neg_excited_before # positive 


# inspired
# before:
neg_inspired_before <- posteriors_PA %>% 
  filter(parameter == "inspired_before") %>% 
  select(posterior) 

# after:
neg_inspired_after <- posteriors_PA %>% 
  filter(parameter == "inspired_after") %>% 
  select(posterior) 

diff_inspired <- neg_inspired_after - neg_inspired_before # positive 



# interested
# before:
neg_interested_before <- posteriors_PA %>% 
  filter(parameter == "interested_before") %>% 
  select(posterior) 

# after:
neg_interested_after <- posteriors_PA %>% 
  filter(parameter == "interested_after") %>% 
  select(posterior) 

diff_interested <- neg_interested_after - neg_interested_before # positive 


# proud
# before:
neg_proud_before <- posteriors_PA %>% 
  filter(parameter == "proud_before") %>% 
  select(posterior) 

# after:
neg_proud_after <- posteriors_PA %>% 
  filter(parameter == "proud_after") %>% 
  select(posterior) 

diff_proud <- neg_proud_after - neg_proud_before # positive 


# strong
# before:
neg_strong_before <- posteriors_PA %>% 
  filter(parameter == "strong_before") %>% 
  select(posterior) 

# after:
neg_strong_after <- posteriors_PA %>% 
  filter(parameter == "strong_after") %>% 
  select(posterior) 

diff_strong <- neg_strong_after - neg_strong_before # positive 

```
```{r pa-posterior-descriptions, echo=FALSE}
rope_range <- rope_range(PA_model)

print("active")
describe_posterior(diff_active, rope_range = rope_range)
print("alert")
describe_posterior(diff_alert, rope_range = rope_range)
print("attentive")
describe_posterior(diff_attentive, rope_range = rope_range)
print("determined")
describe_posterior(diff_determined, rope_range = rope_range)
print("enthusiastic")
describe_posterior(diff_enthusiastic, rope_range = rope_range)
print("excited")
describe_posterior(diff_excited, rope_range = rope_range)
print("inspired")
describe_posterior(diff_inspired, rope_range = rope_range)
print("interested")
describe_posterior(diff_interested, rope_range = rope_range)
print("proud")
describe_posterior(diff_proud, rope_range = rope_range)
print("strong")
describe_posterior(diff_strong, rope_range = rope_range)
```
