---
title: "PANAS Gender Model"
author: "Kristina Kobrock"
date: "14 3 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)
```
```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
# package to extract R data from matlab files
library(R.matlab)

# package for convenience functions (e.g. ggplot2, dplyr, etc.)
library(tidyverse)

# package for Bayesian regression modeling
library(brms)

# package for visualization
library(tidybayes)

# package for plotting (together with ggplot)
library(ggpubr)

# package to visualize 
library(bayesplot)

# package to extract CrIs
library(bayestestR)

# package to extract HDIs
library(HDInterval)

#devtools::install_github("michael-franke/aida-package")
library(aida)

# use the aida-theme for plotting
theme_set(theme_aida())

# global color scheme / non-optimized
# Green RYB: 5EB045, Barn red: 811F0E, Charcoal: 424B54, maximum yellow red: FFBA49, copper crayola: DE8F6E
green <- "#5EB045"
red <- "#811F0E"
gray <- "#424B54"
yellow <- "#FFBA49"
copper <- "#DE8F6E"
black <- "#000000"
project_colors = c(red, green, gray, yellow, copper, black)

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
} 

```

# The role of affect in the Automatic Approach Bias: An empirical investigation.

## Mood induction
Load the preprocessed data (see DataAnalysis.Rmd for details).
```{r}
load("panas.Rdata")
```

### Is there a main effect of participants' gender on mood induction?

#### Visual inspection
```{r aggregate, echo=FALSE}
df_affect_agg <- df_affect_mutate %>% 
  group_by(mood_induction, item_affect, subjectgender, panas_nr) %>% 
  summarise(mean = round(mean(panas_score), 2),
            sd = round(sd(panas_score), 2)) %>% 
  filter(panas_nr == 2) %>% 
  print()
```

Visual inspection of the means does not reveal any differences between gender after the mood induction.

### Statistical model
The variables subjectgender and mood_induction are contrast-coded for quick interpretation of interactions.
```{r contrast-code, echo=FALSE}
df_affect_coded <- df_affect_mutate %>% 
  mutate(subjectgender = ifelse(subjectgender == "female", 0.5, -0.5),
         mood_induction = ifelse(mood_induction == "positive", -0.5, 0.5))
head(df_affect_coded)
```
```{r panas_model, echo=FALSE}
# predicting panas_score by the valence of the mood_induction (positive/negative), panas_nr (before or after the mood induction) and subjectgender

priors = priors <- c(
  set_prior("normal(-22, 5)", class = "Intercept"), # prior informed by the data (NA means are between -20.83 and -23.62), but with a wide standard deviation
  set_prior("normal(44, 5)", class = "b", coef = "item_affectPA"), # prior informed by the data (PA means are between 20.47 and 24.52)
  set_prior("normal(0, 10)", class = "b") # agnostic priors for all other fixed effects
)
# item_affect is not of further interest here, but needs to be included in the model so that it can estimate the correct PA/NA scales
panas_model <- brm(panas_score ~ subjectgender * item_affect * panas_nr * mood_induction, # four-way interaction
                   prior = priors, # same priors as in main model
                   iter = 2000,
                   seed = 305,
                   data = df_affect_coded,
                   file = "panas_gender_model") 
```
```{r}
summary(panas_model, priors=TRUE)
```
```{r pp-check, echo=FALSE}
pp_check(panas_model, ndraws = 50, type = "dens_overlay")
```

There is no main effect of subjectgender on the mood induction. Also none of the interactions including subjectgender substantially contributed to the effect. If anything, then there is a slight difference in baseline mood before mood induction between the two genders (subjectgender:item_affectPA, subjectgender:item_affectPA:mood_induction) - but that's not of importance here, so it's not further investigated.