---
title: "Stimulus Valence"
author: "Kristina Kobrock"
date: "14 3 2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)
```
```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
# package to extract R data from matlab files
library(R.matlab)

# package for convenience functions (e.g. ggplot2, dplyr, etc.)
library(tidyverse)

# package for Bayesian regression modeling
library(brms)

# package for visualization
library(tidybayes)

# package for plotting (together with ggplot)
library(ggpubr)

# package to visualize 
library(bayesplot)

# package to extract CrIs
library(bayestestR)

# package to extract HDIs
library(HDInterval)

#devtools::install_github("michael-franke/aida-package")
library(aida)

# use the aida-theme for plotting
theme_set(theme_aida())

# global color scheme / non-optimized
# Green RYB: 5EB045, Barn red: 811F0E, Charcoal: 424B54, maximum yellow red: FFBA49, copper crayola: DE8F6E
green <- "#5EB045"
red <- "#811F0E"
gray <- "#424B54"
yellow <- "#FFBA49"
copper <- "#DE8F6E"
black <- "#000000"
project_colors = c(red, green, gray, yellow, copper, black)

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
} 

```

# The role of affect in the Automatic Approach Bias: An empirical investigation.

## The effect of stimulus valence in the Affect-AAT
Load the preprocessed data (see DataAnalysis.Rmd for details).
```{r load-data}
load("affect-aat.Rdata")
```

```{r backtransform-function, echo=FALSE}
backtransform_RT <- function(list, df=df_aat_log1) {
  round(exp(list * attr(df$log_RT_s, 'scaled:scale') + attr(df$log_RT_s, 'scaled:center')),2)
}
```

### Model with stimulus valence
```{r data-log-model, echo=FALSE}
df_aat_log1 <- df_aat_log %>% 
  mutate(mood_induction = factor(ifelse(mood_induction == "p", "positive", "negative")),
         condition = factor(condition)) %>% 
  select(subject_id, trial, condition, mood_induction, log_RT_s, pictures, stimulus_valence)
head(df_aat_log1)
```
The same priors are used as in the main analysis. No group-level effects are added.
```{r priors, echo=FALSE}
priors <- c(set_prior("student_t(3, 0, 3)", class = "Intercept"), # model-specific parameters weakly informed by the data
                      set_prior("student_t(3, 0, 3)", class = "sigma"),
                      set_prior("normal(0, 10)", class = "b")
                      ) 
```

One simple model where stimulus_valence is added without interaction:
```{r stimulus-model-no-interaction, echo=FALSE}
fit_stimulusModel <- brm(log_RT_s ~ condition * mood_induction + stimulus_valence,
                         data = df_aat_log1,
                         family = gaussian(),
                         prior = priors,     
                         iter = 2000, 
                         seed = 228,
                         file = "simple_stimulus_model"
                         )
```
```{r}
summary(fit_stimulusModel, priors=TRUE)
```
The model output suggests something that I did not expect: a main effect of stimulus_valence, but a negative main effect suggesting that actually positive instead of negative reaction times are shorter.

I'm running a model with interaction and group-level effects to see whether this main effect is influenced by a specific condition in the experiment:
```{r constrpriors, echo=FALSE}
constr_priors <- c(set_prior("student_t(5, -0.2, 1)", class = "Intercept"), # constraining prior informed by previous model fit and data: centered on -0.2, ranges between -5 and 5, most values being closer to mu
            set_prior("student_t(3, 0, 2)", class = "sigma"),
            set_prior("normal(0, 10)", class = "b"), # agnostic prior on fixed effects
            set_prior("student_t(3, 0, 2)", class = "sd") # weakly informative prior on random effects
            ) 
```
```{r stimulus-model-RE, echo=FALSE}

fit_stimulusModelInt <- brm(log_RT_s ~ condition * mood_induction * stimulus_valence
                           + (condition * mood_induction || subject_id),
                           data = df_aat_log1,
                           family = gaussian(),
                           prior = constr_priors,
                           cores = getOption("mc.cores", 2),
                           iter = 10000, 
                           seed = 1703,
                           file = "stimulus_model"
                           )
```
```{r log-model-summary, echo=FALSE}
summary(fit_stimulusModelInt, priors = TRUE)
```
The main effect of stimulus valence is still visible, reaction times in response to positive stimuli seem to be shorter than reaction times in response to negative stimuli. 

Plot data again to see whether this effect is also visible in the data:
```{r plot-data-agg}
# creating an aggregated data frame grouped by mood_induction and condition with the respective means
df_aat_agg1 <- df_aat_log1 %>% 
  group_by(stimulus_valence, condition) %>% 
  summarise(mean = mean(log_RT_s)) 
```
The plot does not show much difference between groups. What happens if we zoom in?
```{r plot-closer-stimulus-valence, echo=FALSE}
# plot closer
df_aat_log1 %>% 
  # plotting both mood_induction and condition using facet_wrap
  ggplot(aes(x = stimulus_valence, y = log_RT_s, color = stimulus_valence)) +
  facet_wrap(~ condition) +
  geom_point(alpha = 0.1, size = 1,
             # adding a little bit of jitter to make the points visible
             position = position_jitter(width = 0.3)) +
  # adding means and a connecting line in order to make the differences between locations and conditions 
  # visible
  geom_line(data = df_aat_agg1,
            aes(x = stimulus_valence, y = mean, group = condition), color=gray) +
  geom_point(data = df_aat_agg1,
             aes(x = stimulus_valence, y = mean, fill = stimulus_valence), size = 6) +
  # axes labels and titles
  labs(
    title = "Effects of stimulus valence and condition on RT",
    subtitle = "zoomed in",
    y = "log(RT) (scaled)",
    x = "stimulus valence by condition"
  ) +
  guides(fill = "none", color = "none") +
  ylim(-0.2, 0.2)
```

```{r extract-posteriors-cond, echo=FALSE}
posteriors_stimulusModel_cond <- fit_stimulusModelInt %>%
  # just taking the population level effects with regards to stimulus valence and condition here
  spread_draws(b_Intercept, b_conditionincongruent,
               #b_mood_inductionpositive, 
               b_stimulus_valencepositive,
               #`b_conditionincongruent:mood_inductionpositive`,
               `b_conditionincongruent:stimulus_valencepositive`#,
               #`b_mood_inductionpositive:stimulus_valencepositive`,
               #`b_conditionincongruent:mood_inductionpositive:stimulus_valencepositive`
               ) %>% 
  mutate(
    congruent_negative = b_Intercept,
    incongruent_negative = b_Intercept + b_conditionincongruent,
    congruent_positive = b_Intercept + b_stimulus_valencepositive,
    incongruent_positive = b_Intercept + b_conditionincongruent + 
      b_stimulus_valencepositive + `b_conditionincongruent:stimulus_valencepositive` 
    ) %>% 
  dplyr::select(congruent_negative, incongruent_negative,
                congruent_positive, incongruent_positive) %>%
  gather(key = "parameter", value = "posterior")

```

```{r differences-cond, echo=FALSE}

# difference between congruent and incongruent condition regardless of mood induction:

# negative stimulus valences:
post_negative <- posteriors_stimulusModel_cond %>% 
  filter(parameter == "congruent_negative" | parameter == "incongruent_negative") %>% 
  select(posterior) 

# positive stimulus valences:
post_positive <- posteriors_stimulusModel_cond %>% 
  filter(parameter == "congruent_positive" | parameter == "incongruent_positive") %>% 
  select(posterior) 

diff_valence <- post_positive - post_negative # positive when RTs in response to negative stimuli are shorter (as would be expected)


# congruent condition:
post_congruent_neg <- posteriors_stimulusModel_cond %>% 
  filter(parameter == "congruent_negative") %>% 
  select(posterior) 

post_congruent_pos <- posteriors_stimulusModel_cond %>% 
  filter(parameter == "congruent_positive") %>% 
  select(posterior)

diff_valence_congruent <- post_congruent_pos - post_congruent_neg # positive when RTs in response to negative stimuli are shorter (as would be expected)


# incongruent condition:
post_incongruent_neg <- posteriors_stimulusModel_cond %>% 
  filter(parameter == "incongruent_negative") %>% 
  select(posterior)

post_incongruent_pos <- posteriors_stimulusModel_cond %>% 
  filter(parameter == "incongruent_positive") %>% 
  select(posterior) 

diff_valence_incongruent <- post_incongruent_pos - post_incongruent_neg 

```
```{r plot-stimulus-condition-difference, echo=FALSE}
plot_1 <- diff_valence %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = yellow, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "stimulus valence posterior difference"
  ) #+
  #ylim(0, 13)

plot_2 <- diff_valence_congruent %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = yellow, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "congruent trials"
  ) #+
  #ylim(0, 13)

plot_3 <- diff_valence_incongruent %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = yellow, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "incongruent trials"
  ) #+
  #ylim(0, 15)

ggarrange(plot_1, ggarrange(plot_2, plot_3, ncol=2), nrow=2)
```

The difference in reaction times does not seem to depend on condition. 
```{r extract-posterior-mi, echo=FALSE}
posteriors_stimulusModelMI <- fit_stimulusModelInt %>%
  # just taking the population level effects here
  spread_draws(b_Intercept, 
               #b_conditionincongruent,
               b_mood_inductionpositive, 
               b_stimulus_valencepositive,
               `b_conditionincongruent:mood_inductionpositive`,
               #`b_conditionincongruent:stimulus_valencepositive`#,
               `b_mood_inductionpositive:stimulus_valencepositive`#,
               #`b_conditionincongruent:mood_inductionpositive:stimulus_valencepositive`
               ) %>% 
  mutate(
    negMI_negative = b_Intercept,
    posMI_negative = b_Intercept + b_mood_inductionpositive,
    negMI_positive = b_Intercept + b_stimulus_valencepositive,
    posMI_positive = b_Intercept + b_mood_inductionpositive + 
      b_stimulus_valencepositive + `b_mood_inductionpositive:stimulus_valencepositive` 
    ) %>% 
  dplyr::select(negMI_negative, posMI_negative,
                negMI_positive, posMI_positive) %>%
  gather(key = "parameter", value = "posterior")

```

```{r differences-mi, echo=FALSE}

# difference between valence and mood induction regardless of condition:

# negative stimulus valences:
post_negative <- posteriors_stimulusModelMI %>% 
  filter(parameter == "posMI_negative" | parameter == "negMI_negative") %>% 
  select(posterior) 

# positive stimulus valences:
post_positive <- posteriors_stimulusModelMI %>% 
  filter(parameter == "posMI_positive" | parameter == "negMI_positive") %>% 
  select(posterior) 

diff_valence <- post_positive - post_negative # positive when RTs in response to negative stimuli are shorter (as would be expected)


# after negative mood induction:
post_negMI_neg <- posteriors_stimulusModelMI %>% 
  filter(parameter == "negMI_negative") %>% 
  select(posterior) 

post_negMI_pos <- posteriors_stimulusModelMI %>% 
  filter(parameter == "negMI_positive") %>% 
  select(posterior)

diff_valence_negMI <- post_negMI_pos - post_negMI_neg # positive when RTs in response to negative stimuli are shorter (as would be expected)


# after positive mood induction:
post_posMI_neg <- posteriors_stimulusModelMI %>% 
  filter(parameter == "posMI_negative") %>% 
  select(posterior)

post_posMI_pos <- posteriors_stimulusModelMI %>% 
  filter(parameter == "posMI_positive") %>% 
  select(posterior) 

diff_valence_posMI <- post_posMI_pos - post_posMI_neg 

```
```{r plot-stimulus-MI-difference, echo=FALSE}
plot_1 <- diff_valence %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = yellow, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "stimulus valence posterior difference"
  ) #+
  #ylim(0, 13)

plot_2 <- diff_valence_negMI %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = red, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "after negative mood ind."
  ) +
  #ylim(0, 13)
  xlim(-0.25, 0.05)

plot_3 <- diff_valence_posMI %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = green, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "after positive mood ind."
  ) +
  #ylim(0, 15)
  xlim(-0.25, 0.05)
  
ggarrange(plot_1, ggarrange(plot_2, plot_3, ncol=2), nrow=2) +
  ggtitle("Stimulus model: Posterior difference distributions\n") +
  theme(plot.title = element_text(hjust = 0.5, face="bold", size=20))
```
```{r difference-in-ms, echo=FALSE}
# transform all posterior estimates to the ms scale:
post_positive_ms <- backtransform_RT(post_positive)
post_negative_ms <- backtransform_RT(post_negative)
diff_valence_ms <- post_positive_ms - post_negative_ms

post_positive_negMI_ms <- backtransform_RT(post_negMI_pos)
post_negative_negMI_ms <- backtransform_RT(post_negMI_neg)
diff_valence_negMI_ms <- post_positive_negMI_ms - post_negative_negMI_ms

post_positive_posMI_ms <- backtransform_RT(post_posMI_pos)
post_negative_posMI_ms <- backtransform_RT(post_posMI_neg)
diff_valence_posMI_ms <- post_positive_posMI_ms - post_negative_posMI_ms

```
```{r describe-differences-ms, echo=FALSE}

df_diffvalence <- data.frame(
  posteriors = "valence difference",
  median = median(diff_valence_ms$posterior),
  CrI_lower = HDInterval::hdi(diff_valence_ms$posterior)[1],
  CrI_upper = HDInterval::hdi(diff_valence_ms$posterior)[2]
)
df_diffnegMI <- data.frame(
  posteriors = "difference after negative mood induction",
  median = median(diff_valence_negMI_ms$posterior),
  CrI_lower = HDInterval::hdi(diff_valence_negMI_ms$posterior)[1],
  CrI_upper = HDInterval::hdi(diff_valence_negMI_ms$posterior)[2]
)
df_diffposMI <- data.frame(
  posteriors = "difference after positive mood induction",
  median = median(diff_valence_posMI_ms$posterior),
  CrI_lower = HDInterval::hdi(diff_valence_posMI_ms$posterior)[1],
  CrI_upper = HDInterval::hdi(diff_valence_posMI_ms$posterior)[2]
)

differences_in_ms <- tibble(rbind(
  df_diffvalence,
  df_diffnegMI,
  df_diffposMI
))

differences_in_ms
```

The difference seems to depend on the mood induction condition. Especially in the positive mood induction condition, the reaction times are shorter for positive than for negative stimuli.
```{r bayestests, echo=FALSE}
stimulusModel_rope_range = rope_range(fit_stimulusModelInt) # calculates the rope range using the tenth (1/10 = 0.1) of the standard deviation (SD) of the response variable (panas_score), which can be considered as a “negligible” effect size (Cohen, 1988)

print("valence difference regardless of condition:")
posterior_diffcondition <- describe_posterior(diff_valence, centrality = "median", ci = 0.95, ci_method = "hdi", rope_range = stimulusModel_rope_range, rope_ci = 0.95) %>% print()

print("valence difference after negative mood induction:")
posterior_diffv_negMI <- describe_posterior(diff_valence_negMI, rope_range = stimulusModel_rope_range) %>% print()

print("valence difference after positive mood induction:")
posterior_diffv_posMI <- describe_posterior(diff_valence_posMI, rope_range = stimulusModel_rope_range) %>% print()

```
```{r}
describe_posterior(fit_stimulusModelInt, centrality = "median", ci = 0.95, ci_method = "hdi", test = c("p_direction"))
```

```{r plot-data-agg2, echo=FALSE}
# creating an aggregated data frame grouped by mood_induction and condition with the respective means
df_aat_agg2 <- df_aat_log1 %>% 
  group_by(stimulus_valence, mood_induction) %>% 
  summarise(mean = mean(log_RT_s)) 
```
```{r plot-closer-stimulus-valence2, echo=FALSE}
# plot closer
df_aat_log1 %>% 
  # plotting both mood_induction and condition using facet_wrap
  ggplot(aes(x = stimulus_valence, y = log_RT_s, color = stimulus_valence)) +
  facet_wrap(~ mood_induction) +
  geom_point(alpha = 0.1, size = 1,
             # adding a little bit of jitter to make the points visible
             position = position_jitter(width = 0.3)) +
  # adding means and a connecting line in order to make the differences between locations and conditions 
  # visible
  geom_line(data = df_aat_agg2,
            aes(x = stimulus_valence, y = mean, group = mood_induction), color=gray) +
  geom_point(data = df_aat_agg2,
             aes(x = stimulus_valence, y = mean, fill = stimulus_valence), size = 6) +
  # axes labels and titles
  labs(
    title = "Effects of stimulus valence and condition on RT",
    subtitle = "zoomed in",
    y = "log(RT) (scaled)",
    x = "stimulus valence by mood induction"
  ) +
  guides(fill = "none", color = "none") +
  ylim(-0.2, 0.2)
```
```{r plot-stimulus-cond-mi, echo=FALSE}
plot_cond <- df_aat_log1 %>% 
  # plotting both mood_induction and condition using facet_wrap
  ggplot(aes(x = stimulus_valence, y = log_RT_s, color = stimulus_valence)) +
  facet_wrap(~ condition) +
  geom_point(alpha = 0.1, size = 1,
             # adding a little bit of jitter to make the points visible
             position = position_jitter(width = 0.3)) +
  # adding means and a connecting line in order to make the differences between locations and conditions 
  # visible
  geom_line(data = df_aat_agg1,
            aes(x = stimulus_valence, y = mean, group = condition), color=gray) +
  geom_point(data = df_aat_agg1,
             aes(x = stimulus_valence, y = mean, fill = stimulus_valence), size = 4) +
  # axes labels and titles
  labs(
    y = "log(RT) (scaled)",
    x = "condition"
  ) +
  guides(fill = "none", color = "none") +
  ylim(-0.2, 0.2)

plot_MI <- df_aat_log1 %>% 
  # plotting both mood_induction and condition using facet_wrap
  ggplot(aes(x = stimulus_valence, y = log_RT_s, color = stimulus_valence)) +
  facet_wrap(~ mood_induction) +
  geom_point(alpha = 0.1, size = 1,
             # adding a little bit of jitter to make the points visible
             position = position_jitter(width = 0.3)) +
  # adding means and a connecting line in order to make the differences between locations and conditions 
  # visible
  geom_line(data = df_aat_agg2,
            aes(x = stimulus_valence, y = mean, group = mood_induction), color=gray) +
  geom_point(data = df_aat_agg2,
             aes(x = stimulus_valence, y = mean, fill = stimulus_valence), size = 4) +
  # axes labels and titles
  labs(
    x = "mood induction",
    y = ""
  ) +
  guides(fill = "none", color = "none") +
  ylim(-0.2, 0.2)

ggarrange(plot_cond, plot_MI)
```

```{r plot-stimulus-hdi-posteriors, echo=FALSE}
# Extract the posteriors
posteriors_MI <- posteriors_stimulusModelMI %>% 
  group_by(parameter) %>% 
  summarise(mean_posterior = mean(posterior),
            `95lowerCrI` = HDInterval::hdi(posterior, credMass = 0.95)[1],
            `95higherCrI` = HDInterval::hdi(posterior, credMass = 0.95)[2]) %>% 
  mutate(stimulus_valence = ifelse(parameter == "negMI_negative" | parameter == "posMI_negative",
                            "negative", "positive"),
         mood_induction = ifelse(parameter == "negMI_negative" | parameter == "negMI_positive",
                                 "negative mood ind.", "positive mood ind.")
  ) %>% 
  select(mood_induction, stimulus_valence, mean_posterior, `95lowerCrI`, `95higherCrI`) 

# prepare data frames for plotting:
df_aat_log1_plot <- df_aat_log1 %>% 
  mutate(mood_induction = ifelse(mood_induction == "positive", "positive mood ind.", "negative mood ind.")) %>% 
  arrange(mood_induction)
df_aat_agg2_plot <- df_aat_agg2 %>% 
  mutate(mood_induction = ifelse(mood_induction == "positive", "positive mood ind.", "negative mood ind."))

# plot
ggplot(data = posteriors_MI, 
       aes(x = stimulus_valence, y = mean_posterior,
           color = stimulus_valence)) + 
  geom_point(data = df_aat_log1_plot, 
             aes(x = stimulus_valence, y = log_RT_s), alpha = 0.01, size = 2, 
             # manually color coding for mood_induction:
             color=c(rep(red, 28846/2), rep(green, 28846/2)),
             # adding a little bit of jitter to make the points visible
             position = position_jitter(width = 0.3)) +
  geom_errorbar(aes(ymin = `95lowerCrI`, ymax = `95higherCrI`),
                width = 0.2, color = "gray", size=1) +
  geom_line(aes(group = 1), color=gray) +
  geom_point(size = 5) +
  geom_line(data = df_aat_agg2_plot,
            aes(x = stimulus_valence, y = mean, group = mood_induction), color=gray, linetype = "dashed") +
  geom_point(data = df_aat_agg2_plot,
             aes(x = stimulus_valence, y = mean, fill = stimulus_valence), size = 4) +
  facet_grid(~mood_induction) +
  guides(color = "none", fill = "none") +
  ylim(-0.3, 0.2) +
  labs(x = "stimulus valence by mood induction",
       y = "log(RT) (scaled)",
       title = "Stimulus model post. estimates and data means") +
  theme(plot.title = element_text(hjust = 0.5, size=19))
```


```{r hypothesistests}
# difference between reaction times for the different valences?

# SHORTER! reaction times for positive than for negative stimuli in the congruent condition after Negind
hypothesis(fit_stimulusModelInt, 'Intercept + stimulus_valencepositive < Intercept')

# SHORTER reaction times for positive than for negative in the incongruent condition after Negind
hypothesis(fit_stimulusModelInt, 'Intercept + stimulus_valencepositive + conditionincongruent + conditionincongruent:stimulus_valencepositive < Intercept + conditionincongruent')

# after positive Mood ind:
# congruent: LARGEST ESTIMATE!
hypothesis(fit_stimulusModelInt, 'Intercept + stimulus_valencepositive + mood_inductionpositive + mood_inductionpositive:stimulus_valencepositive < Intercept + mood_inductionpositive')

# incongruent:
hypothesis(fit_stimulusModelInt, 'Intercept + stimulus_valencepositive + mood_inductionpositive + mood_inductionpositive:stimulus_valencepositive + conditionincongruent + conditionincongruent:mood_inductionpositive + conditionincongruent:stimulus_valencepositive + conditionincongruent:mood_inductionpositive:stimulus_valencepositive < Intercept + mood_inductionpositive + conditionincongruent + conditionincongruent:mood_inductionpositive')
```

Usually, I'd expect reaction times in response to negative stimuli to be smaller than reaction times in response to positive stimuli. Here, the exact opposite seems to be the case.

#### Comparing with data from previous experiment:
```{r load-raw-data}
df_AAT_raw = read.csv("rohdaten_johannes.csv")
load("affect-aat-raw-data.Rdata")
```
```{r preprocess-data, echo=FALSE}
# bring raw data from both experiments to the same general form

df_affectAAT <- df_aat_excc %>% 
  mutate(study = "Affect-AAT") %>% 
  select(study, reaction_times, subject_id, condition, stimulus_valence, pictures, correct)

df_AAT <- df_AAT_raw %>% 
  filter(condition == "jc" | condition == "ji") %>%   # only joystick conditions are relevant
  mutate(condition = ifelse(condition == "jc", "congruent", "incongruent"), # data should have the same form as Affect-AAT data
         correct = ifelse(comment == "correct trial", "1", "0"),
         stimulus_valence = ifelse(picNumber <= 44, "positive", "negative"),
         study = "AAT"
         ) %>% 
  rename(reaction_times = reactionTime) %>% 
  rename(pictures = picNumber) %>% 
  rename(subject_id = subjectnumber) %>% 
  select(study, reaction_times, subject_id, condition, stimulus_valence, pictures, correct)
 
df_full <- rbind(df_affectAAT, df_AAT)
```
```{r data-for-plotting, echo=FALSE}
df_plot <- df_full %>% 
  mutate(RT_AAT = ifelse(study == "AAT", reaction_times, NA),
         RT_AffectAAT = ifelse(study == "Affect-AAT", reaction_times, NA))
```
```{r plot-study-stimulus, echo=FALSE}
df_plot %>% 
  ggplot(aes(x = pictures, y = reaction_times, color = study)) +
  scale_color_manual(values = c(yellow, copper)) +
  geom_point() + 
  geom_vline(aes(xintercept = 44.5), color=gray, linetype="dashed") +
  ylim(0,5) +
  labs(
    title = "Comparing raw data from AAT and Affect-AAT",
    y = "raw reaction times [s]", 
    x = "pictures (1-44 are positive, 45-88 are negative)"
  )
```
```{r agg-means, echo=FALSE}
df_agg <- df_plot %>% 
  group_by(study, stimulus_valence) %>% 
  summarise(mean = mean(reaction_times)) 
df_agg
```

