---
title: "Affect-AAT model with 2 blocks only"
author: "Kristina Kobrock"
date: "14 3 2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)
```
```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
# package to extract R data from matlab files
library(R.matlab)

# package for convenience functions (e.g. ggplot2, dplyr, etc.)
library(tidyverse)

# package for Bayesian regression modeling
library(brms)

# package for visualization
library(tidybayes)

# package for plotting (together with ggplot)
library(ggpubr)

# package to visualize 
library(bayesplot)

# package to extract CrIs
library(bayestestR)

# package to extract HDIs
library(HDInterval)

#devtools::install_github("michael-franke/aida-package")
library(aida)

# use the aida-theme for plotting
theme_set(theme_aida())

# global color scheme / non-optimized
# Green RYB: 5EB045, Barn red: 811F0E, Charcoal: 424B54, maximum yellow red: FFBA49, copper crayola: DE8F6E
green <- "#5EB045"
red <- "#811F0E"
gray <- "#424B54"
yellow <- "#FFBA49"
copper <- "#DE8F6E"
black <- "#000000"
project_colors = c(red, green, gray, yellow, copper, black)

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
} 

```
```{r backtransform-function, echo=FALSE}
backtransform_RT <- function(list, df=df_aat_log1) {
  round(exp(list * attr(df$log_RT_s, 'scaled:scale') + attr(df$log_RT_s, 'scaled:center')),2)
}
```

# The role of affect in the Automatic Approach Bias: An empirical investigation.

## The effect of mood induction in the AAB

### How long does a mood induction hold?

### Analysis with two instead of four blocks per session
I'm conducting the same preprocessing steps as in the main analysis.
```{r read_data, echo=FALSE}
files_1stsession = file.path("./data/1stsession",  list.files(path="./data/1stsession", pattern=".mat$"))
files_2ndsession = file.path("./data/2ndsession", list.files(path="./data/2ndsession", pattern=".mat$"))
session_1 <- lapply(files_1stsession, readMat)
session_2 <- lapply(files_2ndsession, readMat)
```
```{r preprocess_function_aat, echo=FALSE}
preprocess_aat <- function(data) {
  
  subject_id <- data$subject.index # unique subject ID
  subjectname <- data$subject.name # subject initials
  subjectage <- data$subject.age 
  subjectgender <- data$subject.gender
  subjecthand <- data$subject.hand # subject handedness, i.e. the dominant hand used for writing
  timestamp <- data$subject.date # time log
  session_nr <- data$subject.session # 1 or 2
  mood_induction <- data$mood.induction # positive (p) or negative (n)

  block_order <- data$category # block order: a=congruent-incongruent, b=incongruent-congruent
  if (block_order == 'a') {
    condition <- c(rep(c("congruent", "incongruent", "congruent", "incongruent"), times=c(44,44,44,44)))  
  }
  else {
    condition <- c(rep(c("incongruent", "congruent", "incongruent", "congruent"), times=c(44,44,44,44)))
  }
  
  # sequences: picture order (number 1-44 are positive, 45-88 are negative pictures)
  sequence_1 <- c()
  sequence_2 <- c()
  for (i in 1:176){
    if ((i %% 2) != 0){
      sequence_1 <- c(sequence_1, data$sequences[[i]])
    }
    else {
      sequence_2 <- c(sequence_2, data$sequences[[i]])
    }
  }
  pushtimes <- data$pushtimes # only push trials
  pulltimes <- data$pulltimes # only pull trials
  alltimes <- data$alltimes # all reaction times


  df <- data.frame(subject_id = rep(as.numeric(as.character(subject_id)), 176),
                   subjectname = rep(subjectname, 176),
                   subjectgender = rep(subjectgender, 176),
                   subjectage = rep(as.numeric(as.character(subjectage)), 176), # convert from factor to numeric vector
                   subjecthand = rep(subjecthand, 176),
                   timestamp = rep(timestamp, 176),
                   date = rep(substr(timestamp, 1, 10), 176),
                   year = rep(as.numeric(as.character(substr(timestamp, 1, 4))), 176), # extract year as double
                   month = rep(as.numeric(as.character(substr(timestamp, 6,7))), 176), # extract month as double
                   day = rep(as.numeric(as.character(substr(timestamp, 9, 10))), 176), # extract day as double
                   session_nr = rep(session_nr, 176),
                   mood_induction = rep(mood_induction, 176),
                   block_order = rep(block_order, 176),
                   trial = c(1:176),
                   condition = condition,
                   pictures = c(sequence_1, sequence_2),
                   response = pushtimes[1:176],
                   pushtimes = pushtimes[1:176],
                   pulltimes = pulltimes[1:176],
                   reaction_times = alltimes[1:176] * 1000) %>% # change scaling to ms
    mutate(stimulus = ifelse(pictures <= 44, "positive", "negative"),
           response = ifelse(response == 0, "pull", "push"),
           correct = ifelse(((condition == "congruent" & stimulus == "positive" & response == "pull") |
                               (condition == "congruent" & stimulus == "negative" & response == "push") |
                               (condition == "incongruent" & stimulus == "positive" & response == "push") |
                               (condition == "incongruent" & stimulus == "negative" & response == "pull")), 1, 0))
}
```
```{r preprocessing, include=FALSE}
df_aat <- data.frame()
for (data_file in session_1) {
  df_single <- data.frame(data_file$subjectdata)[[1]]
  df_aat <- rbind(df_aat, preprocess_aat(df_single)) # binds data from each data file together
}
for (data_file in session_2) {
  df_single <- data.frame(data_file$subjectdata)[[1]]
  df_aat <- rbind(df_aat, preprocess_aat(df_single)) # binds also the second session data to the large dataframe
}
head(df_aat)
```
```{r exclude-subjects, include=FALSE}
# check whether for each subjectindex is data available from two sessions and two mood inductions (and the same block order)
df_aat_check <- df_aat %>% 
  group_by(subject_id, session_nr, mood_induction, block_order) %>% 
  summarize() 

# only keep data when data from both sessions is available and no measurement errors occured
df_aat_excs <- df_aat %>% 
  filter(subject_id != 8     # data from 1st session is missing
         & subject_id != 10  # participant didn't fulfill the criteria
         & subject_id != 16  # 2nd session data is missing
         & subject_id != 24  # 2nd session data is missing
         & subject_id != 33  # measurement error (wrong condition)
         & subject_id != 50) # 2nd session data is missing

df_aat_excs %>% group_by(subject_id, session_nr, mood_induction, block_order) %>% 
  summarize() #%>% 
  #print()

```

The two last blocks are excluded.
```{r exclude-last-blocks, echo=FALSE}
# exclude last two blocks
df_aat_excb <- df_aat_excs %>% 
  filter(trial < 89)
```

```{r exclude-practice, echo=FALSE}
# exclude practice trials (first 4 of each block)
df_aat_exct <- df_aat_excb %>% 
  filter(trial > 4
         & (trial < 45 | trial > 48)
         #& (trial < 89 | trial > 92)
         #& (trial < 133 | trial > 136)
         )

nr_exct <- nrow(df_aat_exct)  #%>% print() # 15200 complete data points
```
After having excluded the practice trials, this still leaves us with 15.200 observations.
```{r exclude-incorrect, include=FALSE}
# exclude trials with incorrect responses
df_aat_excc <- df_aat_exct %>% 
  filter(correct == "1") 
  
nr_excc <- nrow(df_aat_excc) %>% print() # 14659

exc_incorrect <- round(((nr_exct - nr_excc) / nr_exct) * 100, 2) %>% print() # 3.56 %
# --> 3.56 % of the data had to be excluded because the response was not correct
```
```{r exclude-large-small, include=FALSE}
# exclude unreasonably small or large reaction times

# exclude reaction times under 150ms because they are probably artifacts
df_aat_excRT <- df_aat_excc %>% 
  filter(reaction_times > 150)

nr_excRT <- nrow(df_aat_excRT) %>% print() # 14628


# what is a sensible upper bound to reaction times? --> 2 to 4 standard deviations above the mean

# trial outliers
sumstat_reaction_times <- df_aat_excRT %>% 
  summarize(min_RT = round(min(reaction_times), 2),    # 166.58 ms
            max_RT = round(max(reaction_times), 2),    # 65479.71 ms
            mean_RT = round(mean(reaction_times), 2),  # 967.67	ms 
            sd_RT = round(sd(reaction_times), 2)       # 759.19	ms 
            ) %>% 
  print()

mean_RT <- sumstat_reaction_times[3] 	
sd_RT <- sumstat_reaction_times[4] 

cutoff_RT <- as.numeric((mean_RT + (2.5 * sd_RT))) %>% print() # 2.5 standard deviations above mean: 2865.645

df_aat_exc <- df_aat_excRT %>% 
  filter(reaction_times < cutoff_RT)

print(nr_excRT - nrow(df_aat_exc)) # --> 148 rows of data excluded

nr_exc <- nrow(df_aat_exc) %>% print() # 14444

# total amount of data excluded:
total_exc <- round(((nr_exct - nr_exc) / nr_exct) * 100, 2) %>% print() # 4.97 %
# --> 4.97 % of the data were excluded in total (incorrect responses, too short or too long reaction times)

# final sample:
df_aat_exc %>% 
  summarize(min = round(min(reaction_times), 2),   # 166.58
            max = round(max(reaction_times), 2),   # 2861.68 ms
            mean = round(mean(reaction_times), 2), # 922.52
            sd = round(sd(reaction_times), 2)      # 349.6
            ) %>% 
  print()
```
After having excluded all incorrect trials and trials with unreasonably short or large reaction times, I'm left with 14.444 observations. Reaction times range between 155.58ms and 2861.68 (mean=922.52, sd=349.6) which is comparable to the main analysis.
```{r log-transform, echo=FALSE}
# standardize and log-transform reaction times
df_aat_log <- df_aat_exc %>% 
  mutate(log_RT = log(reaction_times))
df_aat_log$log_RT_s <- scale(df_aat_log$log_RT, scale = TRUE)
```
```{r data-log-model, echo=FALSE}
df_aat_log1 <- df_aat_log %>% 
  mutate(mood_induction = factor(ifelse(mood_induction == "p", "positive", "negative")),
         condition = factor(condition)) %>% 
  select(subject_id, trial, condition, mood_induction, log_RT_s, pictures, stimulus, response)
head(df_aat_log1)
```
Log-transformed normal model with same specifications as before (same priors and group-level effects for subject_id) for 4000 iterations:
```{r two-block-model, echo=FALSE}
priors <- c(set_prior("student_t(3, 0, 3)", class = "Intercept"), # model-specific parameters weakly informed by the data
                      set_prior("student_t(3, 0, 3)", class = "sigma"),
                      set_prior("normal(0, 10)", class = "b"), # agnostic prior on fixed effects
                      set_prior("student_t(3, 0, 3)", class = "sd") # weakly informative prior on random effects
                      ) 

fit_2block <- brm(log_RT_s ~ condition * mood_induction 
                    + (condition * mood_induction || subject_id), # I expect individual differences (varying intercepts and slopes) between subjects with regards to both the mood induction and the conditions (but no correlation)
                 data = df_aat_log1,
                 family = gaussian(),
                 prior = priors_log_model,     
                 iter = 4000, 
                 cores = getOption("mc.cores", 2), # specify number according to hardware used (up to nr of chains)
                 seed = 1003,
                 file = "two_block_model"
                 )
```
Notably, the two block model converges more easily when including group-level effects for subject_id:
```{r}
summary(fit_2block, priors=TRUE)
```
Results look similar to the main analysis. The hypothesis tests also confirm that for both mood induction conditions, reaction times are longer in the incongruent than in the congruent condition. Notably, this model converges after fewer iterations.

```{r extract-posteriors, echo=FALSE}
posteriors_2block <- fit_2block %>%
  # just taking the population level effects here
  spread_draws(b_Intercept, b_conditionincongruent,
               b_mood_inductionpositive, `b_conditionincongruent:mood_inductionpositive`) %>% 
  mutate(
    congruent_negative = b_Intercept,
    incongruent_negative = b_Intercept + b_conditionincongruent,
    congruent_positive = b_Intercept + b_mood_inductionpositive,
    incongruent_positive = b_Intercept + b_conditionincongruent + 
      b_mood_inductionpositive + `b_conditionincongruent:mood_inductionpositive` 
    ) %>% 
  dplyr::select(congruent_negative, incongruent_negative,
                congruent_positive, incongruent_positive) %>%
  gather(key = "parameter", value = "posterior")

```
```{r plot-hdi-posteriors-2block, echo=FALSE}
# Extract the posteriors
posteriors <- posteriors_2block %>% 
  group_by(parameter) %>% 
  summarise(mean_posterior = mean(posterior),
            `95lowerCrI` = HDInterval::hdi(posterior, credMass = 0.95)[1],
            `95higherCrI` = HDInterval::hdi(posterior, credMass = 0.95)[2])

posteriors <- posteriors %>% 
  mutate(condition = ifelse(parameter == "congruent_negative" | parameter == "congruent_positive",
                            "congruent", "incongruent"),
         mood_induction = ifelse(parameter == "congruent_negative" | parameter == "incongruent_negative",
                                 "negative", "positive")
  ) %>% 
  select(mood_induction, condition, mean_posterior, `95lowerCrI`, `95higherCrI`) 

# plot
ggplot(data = posteriors, 
       aes(x = condition, y = mean_posterior,
           color = mood_induction, fill = mood_induction)) + 
  geom_errorbar(aes(ymin = `95lowerCrI`, ymax = `95higherCrI`),
                width = 0.2, color = "grey") +
  geom_line(aes(group = 1)) +
  geom_point(size = 5) +
  facet_grid(~mood_induction) +
  guides(color = "none", fill = "none") +
  labs(x = "condition by mood induction",
       y = "posterior log(RT) (scaled)",
       title = "Posterior estimates with credible intervals")
```

Compared to the main analysis, here the rather flat slope indicating the AAB after positive mood induction attracts attention. In light of the two hypotheses from the preregistration, this result would be at least as suprising as the result from the main analysis: Not in the negative mood induction condition as predicted but in the positive mood induction condition, is the AAB leveled out! 
```{r differences, echo=FALSE}

# difference between congruent and incongruent condition regardless of mood induction:

# congruent condition:
post_congruent <- posteriors_2block %>% 
  filter(parameter == "congruent_negative" | parameter == "congruent_positive") %>% 
  select(posterior) 

# incongruent condition:
post_incongruent <- posteriors_2block %>% 
  filter(parameter == "incongruent_negative" | parameter == "incongruent_positive") %>% 
  select(posterior) 

diff_condition <- post_incongruent - post_congruent # positive when RTs in incongruent cond are longer than congruent (as expected)


# negative mood induction:
post_congruent_negMI <- posteriors_2block %>% 
  filter(parameter == "congruent_negative") %>% 
  select(posterior) 

post_incongruent_negMI <- posteriors_2block %>% 
  filter(parameter == "incongruent_negative") %>% 
  select(posterior)

diff_condition_negMI <- post_incongruent_negMI - post_congruent_negMI # positive when RTs in incongruent condition longer

# positive mood induction:
post_congruent_posMI <- posteriors_2block %>% 
  filter(parameter == "congruent_positive") %>% 
  select(posterior)

post_incongruent_posMI <- posteriors_2block %>% 
  filter(parameter == "incongruent_positive") %>% 
  select(posterior) 

diff_condition_posMI <- post_incongruent_posMI - post_congruent_posMI # positive when RTs in incongruent cond are longer than congruent (as expected)

```
```{r plot-2block-difference, echo=FALSE}
plot_1 <- diff_condition %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = yellow, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "condition posterior difference"
  ) +
  ylim(0, 13) +
  xlim(-0.1, 0.25)

plot_2 <- diff_condition_negMI %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = red, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "after negative mood ind."
  ) +
  ylim(0, 13) +
  xlim(-0.1, 0.25)

plot_3 <- diff_condition_posMI %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = green, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "after positive mood ind."
  ) +
  ylim(0, 13) +
  xlim(-0.1, 0.25)

ggarrange(plot_1, ggarrange(plot_2, plot_3, ncol=2), nrow=2)
```

Indeed, the difference between congruent and incongruent condition, i.e. the AAB, is not significant after positive mood induction but only after negative mood induction.
```{r difference-in-ms, echo=FALSE}
# transform all posterior estimates to the ms scale:
post_congruent_ms <- backtransform_RT(post_congruent)
post_incongruent_ms <- backtransform_RT(post_incongruent)
diff_condition_ms <- post_incongruent_ms - post_congruent_ms

post_incongruent_negMI_ms <- backtransform_RT(post_incongruent_negMI)
post_congruent_negMI_ms <- backtransform_RT(post_congruent_negMI)
diff_condition_negMI_ms <- post_incongruent_negMI_ms - post_congruent_negMI_ms

post_incongruent_posMI_ms <- backtransform_RT(post_incongruent_posMI)
post_congruent_posMI_ms <- backtransform_RT(post_congruent_posMI)
diff_condition_posMI_ms <- post_incongruent_posMI_ms - post_congruent_posMI_ms

```
```{r plot-2block-differences-ms, echo=FALSE}
plot_1 <- diff_condition_ms %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = yellow, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "condition posterior difference [ms]"
  ) +
  ylim(0, 0.05)

plot_2 <- diff_condition_negMI_ms %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = red, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "after negative mood ind."
  ) +
  ylim(0, 0.05) +
  xlim(-30, 60)

plot_3 <- diff_condition_posMI_ms %>% 
  ggplot(aes(x = posterior)) +
  geom_density(size=1, color = gray, fill = green, alpha = 0.5) +
  geom_vline(aes(xintercept = median(posterior)), size=1, color=gray) +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[1]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = HDInterval::hdi(posterior)[2]), color=gray, linetype="dashed") +
  geom_vline(aes(xintercept = 0), color=red, linetype="dashed") +
  labs(
    x = "after positive mood ind."
  ) +
  ylim(0, 0.05) +
  xlim(-30, 60)

ggarrange(plot_1, ggarrange(plot_2, plot_3, ncol=2), nrow=2)
```
```{r describe-differences-ms, echo=FALSE}

df_diffcondition <- data.frame(
  posteriors = "condition difference",
  median = median(diff_condition_ms$posterior),
  CrI_lower = HDInterval::hdi(diff_condition_ms$posterior)[1],
  CrI_upper = HDInterval::hdi(diff_condition_ms$posterior)[2]
)
df_diffnegMI <- data.frame(
  posteriors = "difference after negative mood induction",
  median = median(diff_condition_negMI_ms$posterior),
  CrI_lower = HDInterval::hdi(diff_condition_negMI_ms$posterior)[1],
  CrI_upper = HDInterval::hdi(diff_condition_negMI_ms$posterior)[2]
)
df_diffposMI <- data.frame(
  posteriors = "difference after positive mood induction",
  median = median(diff_condition_posMI_ms$posterior),
  CrI_lower = HDInterval::hdi(diff_condition_posMI_ms$posterior)[1],
  CrI_upper = HDInterval::hdi(diff_condition_posMI_ms$posterior)[2]
)

differences_in_ms <- tibble(rbind(
  df_diffcondition,
  df_diffnegMI,
  df_diffposMI
))

differences_in_ms
```

```{r bayestests, echo=FALSE}
twoblock_rope_range = rope_range(fit_2block) # calculates the rope range using the tenth (1/10 = 0.1) of the standard deviation (SD) of the response variable (panas_score), which can be considered as a “negligible” effect size (Cohen, 1988)

print("condition difference regardless of mood induction:")
posterior_diffcondition <- describe_posterior(diff_condition, centrality = "median", ci = 0.95, ci_method = "hdi", rope_range = twoblock_rope_range, rope_ci = 0.95) %>% print()

print("condition difference after negative mood induction:")
posterior_diffc_negMI <- describe_posterior(diff_condition_negMI, rope_range = twoblock_rope_range) %>% print()

print("condition difference after positive mood induction:")
posterior_diffc_posMI <- describe_posterior(diff_condition_posMI, rope_range = twoblock_rope_range) %>% print()

print("Is there a difference between the AAB (i.e. difference in RTs between conditions) after positive compared to after negative MI?")
print("positive AAB in negative AAB:")
posterior_diff_AAB <- describe_posterior(diff_condition_posMI, centrality = "none", rope_range = c(posterior_diffc_negMI$CI_low, posterior_diffc_negMI$CI_high), test = c("rope"))
print(posterior_diff_AAB)

print("negative AAB in positive AAB:")
posterior_diff_AAB2 <- describe_posterior(diff_condition_negMI, centrality = "none", rope_range = c(posterior_diffc_posMI$CI_low, posterior_diffc_posMI$CI_high), test = c("rope"))
print(posterior_diff_AAB2)

```

```{r hypothesis-tests}
# Hypothesis 1: After negative mood induction, reaction times are shorter in the incongruent than in the congruent condition.
hypothesis(fit_2block, "conditionincongruent < 0")
# --> reject hypothesis, test other way:
hypothesis(fit_2block, "conditionincongruent > 0")

# Hypothesis 2: After positive mood induction, reaction times are longer in the incongruent than in the congruent condition.
hypothesis(fit_2block, "conditionincongruent + `conditionincongruent:mood_inductionpositive` > 0")
```
Given that when analyzing only the first two blocks, I don't find an AAB in the positive mood induction condition, but only in the negative mood induction condition, the slight evidence from the main analysis in favour of the assumption that there might be a difference in AAB between the two mood induction conditions, has grown. But, of course, the sample size and effective sample size is also smaller than the sample sizes from the main analysis, so the results should be considered with caution.